Connecting the Popularity Adjusted Block Model to the Generalized Random Dot Product Graph for Community Detection and Parameter Estimation

### 2021-03-02 summary.Rmd
more fixes

**Lemma 5.1**. 
Let $P_n = V_n \Lambda_n V_n^\top$ be defined as in Lemma 5.1. Let 
$A_n \sim \Bernoulli(P_n)$ be the adjacency matrix drawn from $P_n$, and let 
$A_n = \hat{V}_n \hat{\Lambda}_n \hat{V}_n^\top$ be its approximate spectral
decomposition using the $K (K+1) / 2$ most positive and $K (K-1) / 2$ most 
negative eigenvalues and their corresponding eigenvectors.
Without loss of generality, order the rows of $V_n$ and $\hat{V}_n$ by 
community, and let $V_n^{(k)}$ and $\hat{V}_n^{(k)}$ be the $n_k \times K^2$ 
matrix of the rows of the $k$^th^ community in $V_n$ and $\hat{V}_n$ 
respectively. Let $r_n^{(k)} = r(V_n^{(k)})$ and 
$\hat{r}_n^{(k)} = r(\hat{V}_n^{(k)})$ be the inradius of points in $V_n$ and
$\hat{V}_n$ as defined by \citet{soltanolkotabi2012} and 
\citet{pmlr-v28-wang13}. Then

\begin{equation} \label{eq:r-conv}
\hat{r}_n^{(k)} \stackrel{p}{\to} r_n^{(k)}
\end{equation}

*Proof of Lemma 5.1*. 
This falls directly out of Theorem 5 of \citet{rubindelanchy2017statistical}. 
Since $\hat{V}_n^{(k)} \stackrel{p}{\to} V_n^{(k)}$ and $r(\cdot)$ is 
continuous, $r(\hat{V}_n^{(k)}) \stackrel{p}{\to} r(V_n^{(k)})$.

**Lemma 5.2**. 
Let $\mu_n^{(k)} = \mu(V_n^{(k)})$ be the subspace incoherence of $V_n^{(k)}$ 
as defined by \citet{soltanolkotabi2012}. Then 
$\mu_n^{(k)} = 0$ 
with probability 1 $\forall n, k$.

*Proof of Lemma 5.2*. 
Since $V_n^{(k)}$ is orthogonal to $V_n^{(l)}$ $\forall k \neq l$ (Lemma 3.1), 
this falls directly out of Theorem 2.8 of \citet{soltanolkotabi2012}.

**Lemma 5.3**. 
Let $\hat{\mu}_n^{(k)} = \mu(\hat{V}_n^{(k)})$ using the extended definition of
$\mu(\cdot)$ from \citet{pmlr-v28-wang13}. Then

\begin{equation} \label{eq:mu-conv}
\hat{\mu}_n^{(k)} \stackrel{p}{\to} 0
\end{equation}

*Proof of Lemma 5.3*. 
This falls directly out of Theorem 5 of \citet{rubindelanchy2017statistical}. 
Since $\hat{V}_n^{(k)} \stackrel{p}{\to} V_n^{(k)}$ and $\mu(\cdot)$ is 
continuous, $\mu(\hat{V}_n^{(k)}) \stackrel{p}{\to} \mu(V_n^{(k)}) = 0$.

**Lemma 5.4**. 
Let $(v_n^{(i)})^\top$ and $(\hat{v}_n^{(i)})^\top$ be the rows of $V_n$ and
$\hat{V}_n$ respectively. By \citet{rubindelanchy2017statistical},

\begin{equation} \label{eq:delta-conv}
\delta_n = \max_i ||\hat{v}_n^{(i)} - v_n^{(i)}|| \stackrel{p}{\to} 0
\end{equation}

@@

By (\ref{eq:r-conv}) and (\ref{eq:mu-conv}),

\begin{equation} \label{eq:thm5-proof}
r(\hat{V}_n^{(k)}) - \mu(\hat{V}_n^{(k)}) \stackrel{p}{\to} 
r_n^{(k)} - \mu_n^{(k)} = r_n^{(k)} > 0
\end{equation}

@@

Then by Theorem 6 of \citet{pmlr-v28-wang13}, (\ref{eq:delta-conv}) and
(\ref{eq:thm5-proof}) imply that $\forall \epsilon > 0$, $\exists N > 0$ such 
that $\forall n > N$, 
$\Pr(\text{SSC detection property holds for } \hat{V}_n) = 1 - \epsilon$.

@@

where each $\hat{\lambda}^{(kl)}$ is defined as in Algorithm 3, approximates 
the latent positions of $M^{(kl)}$ up to indefinite orthogonal transformation 
by the rate given in Theorem 5 of \citet{rubindelanchy2017statistical}.  
In this case, the indefinite orthogonal transformation $W_*$ in the GRDPG 
result \cite{rubindelanchy2017statistical} is of the form 
$U^\top \hat{U}$. The eigenvalues of $M$ are distinct since the signature 
for this GRDPG is $(1, 1)$, and $U^\top \hat{U}$ is block diagonal, 
resulting in $W_* \stackrel{P}{\to} I$. Therefore, the adjacency spectral embedding of $\hat{M}^{(kl)}$ is 
a direct estimation of the specific latent positions outlined for $M^{(kl)}$, 
up to sign flip.

>>>>

**Definition 4** (Inradius) \cite{soltanolkotabi2012} 
\cite{pmlr-v28-wang13}. 
The inradius of a convex body $\mathcal{P}$, denoted by $r(\mathcal{P})$, is 
defined as the radius of the largest Euclidean ball inscribed in $\mathcal{P}$.

**Lemma 5.1**. 
Let $P_n = V_n \Lambda_n V_n^\top$ be defined as in Lemma 5.1. Let 
$A_n \sim \Bernoulli(P_n)$ be the adjacency matrix drawn from $P_n$, and let 
$A_n = \hat{V}_n \hat{\Lambda}_n \hat{V}_n^\top$ be its approximate spectral
decomposition using the $K (K+1) / 2$ most positive and $K (K-1) / 2$ most 
negative eigenvalues and their corresponding eigenvectors.
Without loss of generality, order the rows of $V_n$ and $\hat{V}_n$ by 
community, and let $V_n^{(k)}$ and $\hat{V}_n^{(k)}$ be the $n_k \times K^2$ 
matrix of the rows of the $k$^th^ community in $V_n$ and $\hat{V}_n$ 
respectively. 
Let $\mu_n^{(k)} = \mu(V_n^{(k)})$ be the subspace incoherence of $V_n^{(k)}$ 
as defined by \citet{soltanolkotabi2012}. Then

\begin{equation} \label{eq:mu-zero}
\mu_n^{(k)} = 0
\end{equation}

with probability 1 $\forall n, k$.

Furthermore, Let $\hat{\mu}_n^{(k)} = \mu(\hat{V}_n^{(k)})$ using the 
extended definition of $\mu(\cdot)$ from \citet{pmlr-v28-wang13}. Then

\begin{equation} \label{eq:mu-conv}
\hat{\mu}_n^{(k)} \stackrel{a.s.}{\to} 0
\end{equation}

*Proof of Lemma 5.1*. 
For the first part, 
since $V_n^{(k)}$ is orthogonal to $V_n^{(l)}$ $\forall k \neq l$ (Lemma 3.1), 
this falls directly out of Theorem 2.8 of \citet{soltanolkotabi2012}.

For the second part, 
this falls directly out of Theorem 5 of \citet{rubindelanchy2017statistical}. 
Since $\hat{V}_n^{(k)} \stackrel{a.s.}{\to} V_n^{(k)}$ and $\mu(\cdot)$ is 
continuous, $\mu(\hat{V}_n^{(k)}) \stackrel{a.s.}{\to} \mu(V_n^{(k)}) = 0$.

**Lemma 5.2**. 
Let $(v_n^{(i)})^\top$ and $(\hat{v}_n^{(i)})^\top$ be the rows of $V_n$ and
$\hat{V}_n$ respectively. By \citet{rubindelanchy2017statistical},

\begin{equation} \label{eq:delta-conv}
\delta_n = \max_i ||\hat{v}_n^{(i)} - v_n^{(i)}|| \stackrel{a.s.}{\to} 0
\end{equation}

@@

By (\ref{eq:mu-zero}) and (\ref{eq:mu-conv}),

\begin{equation} \label{eq:thm5-proof}
r(\hat{V}_n^{(k)}) - \mu(\hat{V}_n^{(k)}) \stackrel{a.s.}{\to} 
r(\hat{V}_n^{(k)}) > 0
\end{equation}

@@

Then Theorem 6 of \citet{pmlr-v28-wang13}, (\ref{eq:delta-conv}) and
(\ref{eq:thm5-proof}) imply that the SSC detection property holds as 
$n \to infty$ with probability 1.

@@

where each $\hat{\lambda}^{(kl)}$ is defined as in Algorithm 3, approximates 
the latent positions of $M^{(kl)}$ up to indefinite orthogonal transformation 
by the rate given in Theorem 5 of \citet{rubindelanchy2017statistical}.  
In this case, the indefinite orthogonal transformation $W_*$ in the GRDPG 
result \cite{rubindelanchy2017statistical} is of the form 
$U^\top \hat{U}$. The eigenvalues of $M$ are distinct since the signature 
for this GRDPG is $(1, 1)$, and $U^\top \hat{U}$ is block diagonal, 
resulting in $W_* \stackrel{a.s.}{\to} I$. Therefore, the adjacency spectral
embedding of $\hat{M}^{(kl)}$ is a direct estimation of the specific latent 
positions outlined for $M^{(kl)}$, up to sign flip.
